\section{Hashing}
\begin{multicols*}{2}
\begin{verbatim}
insert(Key k, Value v)
search(Key k)
delete(Key k)
\end{verbatim}
\begin{verbatim}
contains(Key k)
size()
\end{verbatim}
\end{multicols*}
\paragraph{Direct Access Table}
Using an array indexed by keys to access values. Suppose English words, with
max. 28 letters, and each letter represented in 5 bits. Any word can then be
represented in 140 bits. We require a $2^{140}$ sized direct-access array.
\paragraph{Hash Functions $h:U\mapsto\{1..m\}$}
Considers the smaller set of actual keys $K$, and map the
$|K|=n$ keys to $m\approx n$ buckets. By PHP, $\exists h(k_{1})=h(k_{2}) \wedge k_{1}\neq k_{2}$.\\
\textcolor{Blue}{Simple Uniform Hashing Assumption}: Every key is equally
likely to map to every bucket, and keys are mapped independently.\\
Let $X_i=$ numBalls in the $i$-th bucket. $X_{i}\sim Bin(n, \frac{1}{m})$\\
For fixed load $\alpha$, $E($size of max bucket$)=\Theta(\frac{\log n}{\log \log n})$\\
\verb|int Object::hashCode()|: Must redefine \verb|equals|. If two objects are
equal, they have to return the same hash code.\\
Used: \verb|o.hashCode() ^ (table.size() - 1)|
\paragraph{Dealing with Hash Collisions}
\begin{itemize}
  \item New Hash Function\\
    ``Better''? Need to copy table, and eventually another collision
  \item Chaining
  \item Open Addressing
\end{itemize}
\emph{Chaining}: Each bucket contains a linked list\\
\textcolor{Bittersweet}{Space}: $O(m\text{ buckets} + n\text{ items})$\\
\textcolor{Bittersweet}{Runtime}: Insert $O(1 + \verb|cost|(h))$ Delete $O(n + \verb|cost|(h))$
\emph{Open Addressing}: Probing a sequence of buckets until empty found\\
\textcolor{Bittersweet}{Space}: $O(m)$\ \textcolor{Bittersweet}{Runtime}: Worst case $O(n)$ Practice $O(1)$ for
constant load factor and good hash functions
\paragraph{Table Resizing}
\begin{enumerate}
  \item Choose new table size $m'$
  \item Choose new hash function $h'$ such that\\
    $h:U\mapsto\{1..m\}, h':U\mapsto\{1..m'\}$
  \item $\forall{x}\in$ currTable, newTable$[h' (x)] = x$
\end{enumerate}
Resizing costs $\Theta(m' + m + n)$ for new table size $m'$, current table size $m$, $n$ elements.\\
Proof: Consider sized $2n$ table of $n$ elements, and a sequence of $n$
operations. $n$ inserts or deletions are both $\Theta(n)$.
\paragraph{Cuckoo Hashing}
\begin{verbatim}
insert(Key k, Table T, Hash h)
   slot = h(k)
   displaced = T[slot]
   T[slot] = k
   if (displaced != null) then
      insert(displaced,
             T == A ? B : A,
             T == A ? g : f)
\end{verbatim}
