\section{MST}
\paragraph{Properties of MST}
Suppose a connected graph. Otherwise, no spanning tree exists.
\begin{itemize}
  \item MST is a acyclic graph. Tree.
  \item If an edge is removed from a MST, the two components are MSTs.
  \item For every \emph{cycle}, the maximum weighted edge is not in the MST.\@
  \item For every \emph{cut} of the nodes, the minimum weighted edge across the
    cut is in the MST.\@
\end{itemize}
\paragraph{Edge colouring for proof of correctness}
\textcolor{Bittersweet}{Red Rule}: If $C$ is a cycle with no red arcs, then the max-weighted edge in $C$ is coloured red.
\textcolor{Blue}{Blue Rule}: If $D$ is a cut with no blue arcs, then the min-weighted edge in $D$ is coloured blue.
\paragraph{UFDS}
Weighted Union: To make the smaller tree the subtree of the larger tree
$\Rightarrow O(\log(n))$ height.\@ \textcolor{Bittersweet}{Runtime}: $O(\log n)$\\
Path Compression: Shrinks the tree whenever \verb|find| is called.\\
\textcolor{Bittersweet}{Runtime}: $m$ operations on $n$ objects–$O(n +
m\alpha(m, n))$
\begin{verbatim}
findRoot(int p)
   root = p;
   while (parent[root] != root)
      root = parent[root]
   while (parent[p] != p)
      temp = parent[p];
      parent[p] = root;
      p = temp;
   return root;
\end{verbatim}
\paragraph{Kruskal's Algorithm}
\textcolor{Bittersweet}{Runtime}: $O(E\alpha(E)) + O(E\log E)=O(E\log V)$
\begin{itemize}
  \item Initialise UFDS for n nodes
  \item Sort edges by weights
  \item For each edge $e = (u,v)$:
    \begin{itemize}
      \item If $u,v$ are in the same component, skip
      \item Otherwise, add $e$, and union $u, v$
    \end{itemize}
\end{itemize}
\paragraph{Prim's Algorithm}
\textcolor{Bittersweet}{Runtime}: $O((E+V)\log V)$. $O(E + V\log V)$ for fibo-heap.\\
Basic idea:
\begin{itemize}
  \item $S$: set of nodes connected by blue edges
  \item Initially, $S=\{A\}$
  \item Repeat:
    \begin{itemize}
      \item Identify cut: $\{S, V\backslash S\}$
      \item Find minimum weighted edge on cut
      \item Add new node to $S$
    \end{itemize}
\end{itemize}
\paragraph{Variants}
Undirected graphs / Equal weighted edges: DFS / BFS!\\
DAGs: $\forall v\in V$ add minimum weighted incoming edge–$O(E)$\\
Edges weighted $\{1..10\}\Rightarrow$ use an array of size 10 to act as PQ\\
Reweighting: Only relative edge weights matter. Addition, Multiplication
allowed. To find MaxST, multiply by -1 and run MST.\@
